D3064R0: Bringing the Laws of Physics, Computer-System Constraints, and Compilers to Bear on the OOTA Problem

### 2024-02-14 ootavacuous.tex
Provide volatile non-synchronization example
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

Taking our cue from the folklore, we propose to recognize formally
that programs with volatile objects can execute in two different kinds
of environment: a benign one in which accesses to these objects work
the same as non-volatile memory accesses, and a non-benign one in which
accesses to volatile objects are subject to outside interference and
act more like I/O.
In particular, when running in a non-benign environment a program's
volatile loads can return unpredictable values.
They don't necessarily read from stores (in contrast to non-volatile loads,
which always must return the value of the store they read from).
This implies that volatile load-acquire's do not synchronize with
volatile store-releases in the sense of the C++ memory model,
so they do not contribute in the same way to the happens-before relation.
Also, in these environments the rfe relation does not apply
to volatile loads and hence the accesses in an OOTA cycle must be
non-volatile.

>>>>

Taking our cue from the folklore, we propose to recognize formally
that programs with volatile objects can execute in two different kinds
of environment: a benign one in which accesses to these objects work
the same as non-volatile memory accesses, and a non-benign one in which
accesses to volatile objects are subject to outside interference and
act more like I/O.
In particular, when running in a non-benign environment a program's
volatile loads can return unpredictable values.
They don't necessarily read from stores (in contrast to non-volatile loads,
which always must return the value of the store they read from).
This implies that volatile load-acquire's do not necessarily synchronize with
volatile store-releases in the sense of the C++ memory model,\footnote{
	For example, roughly speaking, they might instead synchronize
	with load-acquires and store-releases in device firmware.}
so they do not contribute in the same way to the happens-before relation.
Also, in these environments the rfe relation does not necessarily apply
to volatile loads and hence the accesses in an OOTA cycle must be
non-volatile.
